---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eider

<!-- badges: start -->
<!-- badges: end -->

The goal of eider is to ... **TODO**

## Installation

You can install the development version of eider from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("alan-turing-institute/eider")
```

TODO: write some examples

```{r example}
library(eider)
```

## Development

If you are making changes to the library itself, first clone the repository:

```
git clone git@github.com:alan-turing-institute/eider.git
```

You will need to install the `lintr`, `pkgdown`, `devtools` R packages to build documentation, run tests, and lint.
Then, from the repository root, you can use the following commands:

- `make doc` generates all function documentation, and also generates the `README.md` file from `README.rmd`
- `make lint` lints the project directory
- `make test` runs all tests

You can also use [`pre-commit`](https://pre-commit.com/) to run all of these before committing, to ensure that you do not commit incomplete code.
Firstly, install `pre-commit` according to the instructions on the webpage above.
Then run `pre-commit install`.

*What about vignettes?*
Well, building vignettes is slightly more complicated.
You can perform a one-time build from the R console using `pkgdown::build_site()`, but running this every time you edit a file gets tiring quickly.
To automate this, first install the package with `make install`, and install a working version of Python and also [`entr`](https://github.com/eradman/entr) (the latter is available on Homebrew via `brew install entr`).
Then run `make vig`: this will monitor your vignette RMarkdown files, rebuild the vignettes any time they are changed, and launch a HTTP server on port 8000 to view the files.
If you change any library code you will have to run `make install` again before rerunning `make vig`.
