---
title: "Examples: SMR04 data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples: SMR04 data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eider)
library(magrittr)
```

This series of vignettes in the _Gallery_ section aim to demonstrate the functionality of `eider` through examples that are similar to real-life usage.
To do this, we have created a series of randomly generated datasets that are stored with the package.
You can access these datasets using the `eider_example()` function, which will return the path to where the dataset is stored in your installation of R.

```{r}
smr04_data_filepath <- eider_example("random_smr04_data.csv")

smr04_data_filepath
```

## The data

In this specific vignette, we are using simulated [SMR04 data](https://publichealthscotland.scot/services/national-data-catalogue/smr-data-manual/definitions-by-smr-record-section/smr04-mental-health-inpatient-and-day-case/general-definitions/).
Our dataset does not contain every column specified in here, but serves as a useful example of how real-life data may be treated using `eider`.

```{r}
smr04_data <- utils::read.csv(smr04_data_filepath) %>%
  dplyr::mutate(
    admission_date = lubridate::ymd(admission_date),
    discharge_date = lubridate::ymd(discharge_date)
  )

dplyr::glimpse(smr04_data)
```

(Note that when the data is loaded by `eider`, the date columns are automatically converted to the date type for you: you do not need to do the manual processing above.)

Each row in this table corresponds to one *episode*; multiple episodes may be associated with the same *stay*.
This simplified table has 7 columns:

* `id`, which is a numeric patient ID;
* `admission_date` and `discharge_date`, which are the dates of admission and discharge for each episode;
* `cis_marker`, which is a unique number associated with each stay (note that this is not necessarily unique for different patient IDs);
* `episode_within_cis`, which is the episode number within each stay; and
* `specialty`, which is the specialty code for the episode.


## Feature 1: Number of episodes associated with a psychotherapy specialty

TODO DESCRIBE

```{r}
pt_episodes_filepath <- eider_example("psychotherapy_episodes.json")
writeLines(readLines(pt_episodes_filepath))
```

```{r}
res <- run_pipeline(
  data_sources = list(smr04 = smr04_data_filepath),
  feature_filenames = pt_episodes_filepath
)

dplyr::glimpse(res$features)
```

## Feature 2: Number of stays associated with a psychotherapy specialty

TODO DESCRIBE

```{r}
pt_stays_filepath <- eider_example("psychotherapy_stays.json")
writeLines(readLines(pt_stays_filepath))
```

```{r}
res <- run_pipeline(
  data_sources = list(smr04 = smr04_data_filepath),
  feature_filenames = pt_stays_filepath
)

dplyr::glimpse(res$features)
```

## Interlude: Adding data sources on-the-fly

The last three features will be concerned with the number of days spent in hospital.
In principle, this can be easily calculated from the data we already have: by taking the discharge date and subtracting the admission date, we can obtain the length of each episode.

`eider` does not yet possess the functionality to preprocess tables in this way (by adding new columns).
However, the package _does_ allow you to perform this calculation yourself (e.g. using `dplyr`), and then add the new table to the pipeline as a new data source.
Specifically, data sources do not necessarily need to be CSV filenames; they can simply be data frames themselves.

Let's construct this new data frame:

```{r}
smr04_with_days_data <- smr04_data %>%
  dplyr::mutate(days_in_hospital = as.numeric(discharge_date - admission_date))

dplyr::glimpse(smr04_with_days_data)
```

In the subsequent sections, we'll provide this new data frame as a data source to `run_pipeline()`.

## Feature 3: Total number of days spent in hospital

TODO DESCRIBE

```{r}
total_days_filepath <- eider_example("days_in_smr04.json")
writeLines(readLines(total_days_filepath))
```

```{r}
res <- run_pipeline(
  data_sources = list(smr04_with_days = smr04_with_days_data),
  feature_filenames = total_days_filepath
)

dplyr::glimpse(res$features)
```

## Feature 4: Longest stay in hospital

TODO: DESCRIBE

```{r}
longest_stay_filepath <- eider_example("longest_stay.json")
writeLines(readLines(longest_stay_filepath))
```

```{r}
res <- run_pipeline(
  data_sources = list(smr04_with_days = smr04_with_days_data),
  feature_filenames = longest_stay_filepath
)

dplyr::glimpse(res$features)
```

## Feature 5: Total number of days spent in any psychotherapy-related stay

TODO: DESCRIBE

```{r}
pt_days_filepath <- eider_example("days_in_psychotherapy.json")
writeLines(readLines(pt_days_filepath))
```

```{r}
res <- run_pipeline(
  data_sources = list(smr04_with_days = smr04_with_days_data),
  feature_filenames = pt_days_filepath
)

dplyr::glimpse(res$features)
```

## Putting it all together

Just for good measure, let's run the entire pipeline with all five of the features above in one go.

```{r}
res <- run_pipeline(
  data_sources = list(
    smr04 = smr04_data_filepath,
    smr04_with_days = smr04_with_days_data
  ),
  feature_filenames = c(
    pt_episodes_filepath,
    pt_stays_filepath,
    total_days_filepath,
    longest_stay_filepath,
    pt_days_filepath
  )
)

dplyr::glimpse(res$features)
```
