---
title: "Filtering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filtering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eider)
library(magrittr)
```

Filtering relies on a `filter` key in the feature JSON, for which the corresponding value is an object:

```json
{
    "output_feature_name": "..."
    ...,
    "filter": {
       ...
    },
    ...
}
```

This vignette will describe the types of filters available and how they can be expressed in this JSON object.

## Basic filters

The most basic filter would be to accept all rows where the value in a given column meets a certain criterion.
For example, we may only want to look at rows for which the value in the `age` column is greater than 18.
There are three pieces of information here which must be encoded:

- **(string) `column`:** the name of the column to filter on
- **(string) `type`:** the type of filter to apply
- **`value`:** the value to compare against. The type of this must match the type of the column specified.

The `type` key can take one of the following values:

- `"in"`

   In this case, the `value` key can either be a single value or an array of multiple values. This filers for rows where the value in the column matches one of the provided values.

- `"gt"` / `"lt"` / `"gt_eq"` / `"lt_eq"`

  The value in the column must be greater than, less than, greater than or equal to, or less than or equal to the `value` respectively.
  For these filters, only one `value` can be passed.

- `"date_gt"` / `"date_lt"` / `"date_gt_eq"` / `"date_lt_eq"`

  The same as above, but to be used for dates.
  This distinction is mandatory because JSON does not have a native date type, so the `value` must be passed as a string in the format `YYYY-MM-DD`.

Thus, for example, if we assume that the column `age` contains integers, then

```json
{
    "filter": {
        "column": "age",
        "type": "in",
        "value": [18, 19, 20]
    }
}
```

selects for people who are 18, 19, or 20 years old; and likewise if the `date_of_birth` column contains dates, then

```json
{
    "filter": {
        "column": "date_of_birth",
        "type": "date_gt_eq",
        "value": "2000-01-01"
    }
}
```

selects for people born on or after January 1, 2000.

As a reminder, the table to be filtered is passed as the `source_file` key in the feature JSON.
This is described in the [introductory vignette](eider.html).


## Compound filters

Filters may also be combined using the Boolean logic operators AND, OR, and NOT.
Instead of providing the three keys `column`, `type`, and `value`, only `type` is specified, together with `subfilter`, which is itself an object which maps filter names to filter objects.

An example is probably the easiest way to explain this.
If we wanted to select for people who are *not* 18, 19, or 20 years old, we could write:

```json
{
    "filter": {
        "type": "not",
        "subfilter": {
            "age_filter": {
                "column": "age",
                "type": "in",
                "value": [18, 19, 20]
            }
        }
    }
}
```

The filter name, `age_filter`, can be any string you want: it is used only for logging purposes.
Notice how the object that `age_filter` refers to, i.e.,

```json
{
    "column": "age",
    "type": "in",
    "value": [18, 19, 20]
}
```

is itself just a basic filter as has already been described above.

`or` and `and` filters work in exactly the same way, with the sole difference being that to accomplish any meaningful results you would generally want the `subfilter` object to contain two or more filters that work in tandem.
For example, the below filters for 18-year-olds who have a height of over 180 (presumably centimetres; although this depends on the input table!):

```json
{
    "filter": {
        "type": "and",
        "subfilter": {
            "age_filter": {
                "column": "age",
                "type": "in",
                "value": [18]
            },
            "height_filter": {
                "column": "height",
                "type": "gt",
                "value": 180
            }
        }
    }
}
```

Again, `age_filter` and `height_filter` are just names for logging purposes, and themselves refer to basic filters.

Finally, note that the subfilters need not _necessarily_ be basic filters, though: they can be compound filters themselves.
So it is possible to nest filters as deeply as you like to accomplish the desired result.

## An example

TODO
