---
title: "A & E Example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A & E Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
This vignette is a demonstration of the `eider` package, focussing on accident and emergency (A&E) data.

First, we load the package from the local directory.

```{r}
library(eider)
```

---------

# Process pipeline
## 0. Decide what you want to filter for
`eider` is designed to process (Scottish) medical data and collate using a range of filter functions to allow users to easily extract tables for ML pipelines. 

Argubly the first step should be to decide which data you will use, and how you will filter it. `eider` allows users to specify a range of data sources (files/tables) and to define which features they want to extract data via a specification json. `eider` will filter and join the data as per the spec.


## 1. Read in data
In this example we will use some random A&E-type data stored in a csv file, and filter based on 2 categories. The `attendance_type` column and the `diagnosis_1` column. But data can be derived from multiple sources, and joined subsequently.

This step must be done before any feature extraction to avoid having to read the same table multiple times.

Here we are utilising MINISPARRA's example data which is accessed via `eider_examples("random_ae_data.csv")`

```{r}
ae_data_filepath <- eider_example("random_ae_data.csv")
read.csv(ae_data_filepath)
```

A snippet of the table is shown above.

The table has 6 columns, a patient id `id`, a `date` TODO update header at which the admission occured, an `attendance_type`, and three diagnosis categories, `diagnosis_1`, `diagnosis_2`, and `diagnosis_3`

From the first row in this example we can see that patient 2 was admitted on the 16th of August 2016, they had 3 unique diagnoses, 104, 103, and 102. In the second row, the patient with the same id was admitted on the 28th July 2015, again they had 3 unique diagnosis, this time 101, 103, and 104.

## 2. Define a feature to extract
Users then specify which features they would like to extract from which data. A simple specification file is given below.

`source_file` is the name for the corresponding filepath in a named list (in this case the random A&E data). `transformation_type` is the type of transformation which will take place. eider has N different transformation types, X, Y, and Z. 
The user specifies how the data will be grouped via `grouping_column`, which in this case is grouping by patient ID. An (optional) flag for what to do if data is absent is given - here it will be to report 0. `output_feature_name` defines the header for the column in the output dataframe which will contain the results of the specific transformation

```{r}
basic_ae_spec_filepath <- eider_example("spec_basic_ae_data.json")
writeLines(readLines(basic_ae_spec_filepath))
```

The `primary_filter` is AND, and there are two `subfilters`. The first subfilter - `subfilter_1` specifies that the user wants to filter based on the `attendance_category` column of the random A&E-type data, and that the value that they are interested in is 1. That is only return results where the `attendance_category` is 1.

The second subfilter `subfilter_2` filters based on the `diagnosis_1` column and should only return results where `diagnosis_1` is 101.

As the primary filter is AND, the resultant dataframe will be a count, grouped by patient `id` of A&E data with `attendance_category` = 1 AND `diagnosos_1` = 101.


## 3. Extract the feature

Now we can use `eider` to run the transformation pipeline. We specify the path to the data via `all_table_filenames` and the path the the specification via `all_feature_json_filenames`.

We then run the transformation which will produce the resultant dataframe.
```{r}
ae_data_path <- list(ae2 = ae_data_filepath)
# Note the name ae2 is used in the spec example `source_file` above
transform(ae_data_path, basic_ae_spec_filepath)
```


Above is the resultant dataframe. We can see that patient with id 2 had 1 instance where their `diagnosis_1` was 101, whereas patient with id 7 had 5 instances. It is also noted that patients with id 3, 9, 15, and 16 either do not fit the given filtering criteria or had no occurences in the source data. These patient's resultant column value is the `absent_default_value` from the specification.


## 4. Using date filtering
In the above example the returned dataframe spanned all the dates over which the input data csv spanned. However this can be refined in the specification.
```{r}
basic_ae_spec_w_date_filepath <-
  eider_example("spec_example_basic_ae_data_w_date.json")
writeLines(readLines(basic_ae_spec_w_date_filepath))
```
An additonal subfilter has been added. Subfilter_3 is applied to the `time` column and filters based on dates that are greater than the 31st of December 2015. This means an event which happened on new-years-eve 2015 will not be included in the data, but events which occured on new-years-day 2016 would be.
Also note the `absent_default_value` has now been removed, and a more descriptive `output_feature_name` has been given in the specification.

We can again run this transformation pipeline with the same A&E data:

```{r}
transform(ae_data_path, basic_ae_spec_w_date_filepath)
```

We note that patient with id 2 now has no entries in the resultant table. This is expected as noted above patient 2 only had one instance where their `diagnosis_1` was 101, and from inspection of the source data, we can see that it was on the 28th July 2015. 

## 5. Using OR filter

We can also expand the filters to return results if either `diagnosis_1` or `diagnosis_2` is 101.
The specification file would look like this:

```{r}
ae_spec_or_w_date_fp <- eider_example(
  "spec_example_basic_ae_data_or_w_date.json"
)
writeLines(readLines(ae_spec_or_w_date_fp))
```

Subfilter 2 has now been defined as an OR type, and it contains it's own subfilters, 21 and 22. `subfilter_21` is a filter that `diagnosis_1` column is 101, and `subfilter_22` is a filter that `diagnosis_2` column is 101. So the transformation will count an instance if either `diagnosis_1` OR `diagnosis_2` are 101. Subject to the other filters of the `attendence_type` being 1, and the date being later than 31st December 2015.

```{r}
transform(ae_data_path, ae_spec_or_w_date_fp)
```

Here we see patient 14 has 2 instances of either `diagnosis_1` or `diagnosis_2` being 101 when the `attendence_type` is 1, and for dates later than 31st December 2015.

## 6. Logging

Logging is also present throughout the package.
By default, logs are printed to the console, so in order to see the log output here, we need to capture it in a temporary file:

```{r}
library(logger)
tmp <- tempfile()
log_threshold(TRACE)
log_appender(appender_file(tmp))
```

And the first basic example is run again. Reminder this is the specification:
```{r}
writeLines(readLines(basic_ae_spec_filepath))
```

And the transformation pipeline is run again - for completeness here is the result.
```{r}
transform(ae_data_path, basic_ae_spec_filepath)
```

Check the log:
```{r}
readLines(tmp)
```

<<<<<<< HEAD:quarto/a_and_e.qmd

Line 4 is when the primary filter is parsed. It has two subfilters so it is identified as a nested filter. The two subfilters (which themselves do not contain any subfilters) are then parsed, as reported on lines 5 and 6.
Lines 7 -11 give the context of what is being processed for debug purposes including the path to the corresponding specification files (which in this case are in the examples folder of this package).

## 7.Giving a spec as a JSON string
It is also possible to specify a path to the json file. In this case we have the same information
as in `spec_basic_ae_data.json`

```{r}
json_string <- '{"source_file":["ae2"],"transformation_type":["COUNT"],"grouping_columns":["id"],"absent_data_flag":[0],"output_feature_name":["user_defined_name"],"primary_filter":{"type":["AND"],"subfilters":{"subfilter_1":{"column":["attendance_category"],"type":["IN"],"value":[1]},"subfilter_2":{"column":["diagnosis_1"],"type":["IN"],"value":[101]}}}}'
```
Apply the transformation
```{r}
tmp <- tempfile()
log_threshold(TRACE)
log_appender(appender_file(tmp))

tf <- transform(ae_data_path, json_string)
tf
```

```{r}
readLines(tmp)
```
=======
Line 2 is when the primary filter is parsed. It has two subfilters so it is identified as a nested filter. The two subfilters (which themselves do not contain any subfilters) are then parsed, as reported on lines 3 and 4.
Lines 5 - 9 give the context of what is being processed for debug purposes including the path to the corresponding specification files (which in this case are in the examples folder of this package).
>>>>>>> main:eider/vignettes/a_and_e.Rmd

## Glossary
* The following comparison operators can be used for the type of filter. Note these are case insensitive.
    + "in" = `%in%`
    + "lt" = `<`
    + "lt_eq" = `<=`
    + "gt" = `>`
    + "gt_eq" = `>=`
    + "date_in" = `%in%`
    + "date_lt" = `<`
    + "date_lt_eq" = `<=`
    + "date_gt" = `>`
    + "date_gt_eq" = `>=`

  

